### 2018-06-04 doc/submission/reviewer-response.md

---
output: fost::default_manuscript
---

<!-- add manuscript number to cover letter -->

<!--
The editorial staff will be in touch shortly to suggest revisions to your title
and abstract for you to consider. Please wait to hear from them before
submitting your revised manuscript. 
-->

# Dear Dr.

We'd like to thank the editors and the reviewers for their thorough and
insightful review of our manuscript. We have been able to make the vast majority
of the suggested revisions, which we feel has improved the manuscript considerably.
Below we present a point-by-point response to each of the editors' and reviewers' 
comments.

@@

- All three reviewers agreed that the paper had the strength of longitudinal
followup and would be of interest to readers of the JLR but they also noted
that there have been earlier studies with similar outcomes and one reviewer
noted that some of those papers were by the present group but were not
mentioned here. The reviewers also thought that the discussion was too
speculative and the authors need therefore, to moderate their enthusiasm. My
additional comment is that although you quote one paper by Yki-Jarvinen and
colleagues. you did not discuss/reference her more recent papers linking
shorter chain saturated fatty acids in the liver to NAFLD/NASH. You should look
at those papers and add them to the discussion. I look forward to a modified
version.

    *Thank you for this suggestion. We have made additions to the manuscript based on the editor's and reviewers'
    suggestions, including the discussion around short chain SFA and NAFLD by
    Yki-Jarvinen. We believe this has strengthened the paper and hope it
    satisfactorily addresses the editor and reviewers comments.*

@@

    *We thank the reviewer for noticing those omissions. The r values have been
    included in the text.*

# Reviewer 2

1.	It is a frustrating that this is seemingly the third in a series of papers
that is looking at patterns in lipid fractions and association with metabolic
outcomes. Thus, the TG, NEFA and phospholipid fractions all seem to be a 'least
publishable unit'. The authors must integrate the knowledge gained from the
previous analysis of NEFA and phospholipids and discuss some of the seemingly
contradictory findings from the previous studies. If the data were examined,
would the DNL story 'hold up'?

    *We thank the reviewer for these comments. We agree that ideally the results
    of all the fatty acid fractions would have been included together, but given the number of
    fatty acids, the longitudinal visits, the depth and volume of the analyses,
    and the complex biology involved, it simply wasn't possible to contain
    everything in a single journal paper given word count and table/figure limitations. We have included a brief discussion of
    the results from the other fractions in the Discussion section. The 
    complete and integrated discussion and presentation of the results is found
    in the first author's (LWJ) PhD thesis (http://hdl.handle.net/1807/80893). As to a combined analysis,
    we have presented these results at the European Diabetes Epidemiology Group
    Meeting in April, 2018 (https://doi.org/10.6084/m9.figshare.6159293), and the proposed DNL mechanism
    remained clearly associated and had a strong contribution to associations
    with diabetes pathogenesis compared to other fatty acids/compositions.
    A short aside, these currently unpublished findings do not invalidate or 
    undermine the previous work and the current manuscript. The combined analysis uses a powerful
    statistical technique (PLS), but it has a number of major limitations,
    not least of which is it can only handle cross-sectional data and can not
    adjust for potential confounders. The present analysis in this manuscript uses PLS in addition
    to the powerful longitudinal technique GEE.*

### 2018-06-04 doc/manuscript.Rmd
started fixing minor edits

Authors: Luke W. Johnston, PhD (1); Zhen Liu, PhD (1); Ravi Retnakaran, MD
(3,4); Bernard Zinman, MD (3,4); Adria Giacca, PhD (5); Stewart B. Harris, MD (2); 
Richard P. Bazinet, PhD (1); and Anthony J. Hanley, PhD (1,3,6)

@@

(1) Department of Nutritional Sciences, University of Toronto, Toronto, Ontario,
Canada.
(2) Centre for Studies in Family Medicine, University of Western Ontario,
London, Ontario, Canada.
(3) Leadership Sinai Centre for Diabetes, Division of Endocrinology, University
of Toronto, Toronto, ON, Canada.
(4) Lunenfeld Tanenbaum Research Institute, Mount Sinai Hospital, Toronto,
Ontario, Canada.
(5) Department of Physiology, University of Toronto, Toronto, Ontario, Canada.
(6) Dalla Lana School of Public Health, University of Toronto, Toronto, Ontario,
Canada.

@@

- Name: Anthony J. Hanley
- Current address:  
    Department of Nutritional Sciences  
    Faculty of Medicine  
    University of Toronto  
    FitzGerald Building, 150 College Street, Room 341  
    Toronto, ON, Canada, M5S 3E2
- Phone number: 416.978.3616
- Fax number:
- Email: anthony.hanley@utoronto.ca

@@

**Results**: 
In adjusted GEE models, four TGFA (14:0, 16:0, 14:1n-7, 16:1n-7 as mol%) had
strong negative associations with IS while others (e.g. 18:1n-7, 18:1n-9,
20:2n-6, 20:5n-3) had strong positive associations. Few associations were seen
for beta-cell function, except for 16:0 and 18:1n-7. PLS analysis indicated four
TGFA (14:0, 16:0, 14:1n-7, 16:1n-7) that clustered together and strongly related
with lower IS. These four TGFA also correlated highly (r>0.4) with clinically
measured TG.

@@

# Keywords
<!-- up to 10. Five from list, five free-form. Should not appear in title or
running title -->

>>>>

Authors: Luke W. Johnston, PhD (1); Zhen Liu, PhD (1); Ravi Retnakaran, MD
(2,3); Bernard Zinman, MD (2,3); Adria Giacca, PhD (4); Stewart B. Harris, MD (5);
Richard P. Bazinet, PhD (1); and Anthony J. Hanley, PhD (1,2,6)

@@

(1) Department of Nutritional Sciences, University of Toronto, Toronto, Ontario,
Canada.
(2) Leadership Sinai Centre for Diabetes, Division of Endocrinology, University
of Toronto, Toronto, ON, Canada.
(3) Lunenfeld Tanenbaum Research Institute, Mount Sinai Hospital, Toronto,
Ontario, Canada.
(4) Department of Physiology, University of Toronto, Toronto, Ontario, Canada.
(5) Centre for Studies in Family Medicine, University of Western Ontario,
London, Ontario, Canada.
(6) Dalla Lana School of Public Health, University of Toronto, Toronto, Ontario,
Canada.

@@

- Name: Anthony J. Hanley
- Current address:  
    Department of Nutritional Sciences  
    Faculty of Medicine  
    University of Toronto  
    FitzGerald Building, 150 College Street, Room 341  
    Toronto, ON, Canada, M5S 3E2
- Phone number: 416.978.3616
- Email: anthony.hanley@utoronto.ca

@@

**Results**: 
In adjusted GEE models, four TGFA (14:0, 16:0, 14:1n-7, 16:1n-7 as mol%) had
strong negative associations with IS while others (e.g. 18:1n-7, 18:1n-9,
20:2n-6, 20:5n-3) had strong positive associations. Few associations were seen
for beta-cell function, except for 16:0, 18:1n-7, and 20:2n-6. PLS analysis indicated four
TGFA (14:0, 16:0, 14:1n-7, 16:1n-7) that clustered together and strongly related
with lower IS. These four TGFA also correlated highly (r>0.4) with clinically
measured TG.

@@
